\documentclass[slidestop]{beamer}
\usepackage{beamerthemesplit}
\usepackage{graphics}
\usepackage{pstricks}

\graphicspath{{./}}

\title{The delicate disadvantage of Reverse-Engineering}
\author{Luke Kenneth Casson Leighton}


\begin{document}

\frame{
   \begin{center}
    \huge{The delicate disadvantage of Reverse-Engineering}\\
    \vspace{32pt}
    \Large{The consequences of maintaining}\\
    \Large{proprietary hardware}\\
    \Large{Can we do better?}\\
    \vspace{24pt}
    \Large{[proposed for] OFSC 2020}\\
    \vspace{16pt}
    \large{\today}
  \end{center}
}


\frame{\frametitle{Background (about me)}

\vspace{15pt}

 \begin{itemize}
   \item First reverse-engineering was Samba-TNG\\
   		 NTBugTraq, August 1996\\
   		 "Welcome to the SAMBA Domain"\\
   		 3 years later...\vspace{6pt}
   \item 2002: Exchange 5.5, enhancing FreeDCE \\
   		 Copied by an "Open" team that removed all attribution\vspace{6pt}
   \item 2003-2005: Xanadux Project\\
	     9 HTC smartphones reverse-engineered\\
	     Zero income earned.\vspace{6pt}
   \item Lesson learned: everyone else makes money from your work.
  \end{itemize}
}


\frame{\frametitle{How come I could do this but others couldn't?}

\vspace{10pt}

 \begin{itemize}
   \item Self-analysis time: what capability did I have \\
   		 that others do not?\\
   		 \vspace{12pt}
   \item Definition of Reverse-Engineering:\\
   		 \vspace{4pt}
   		 The ability to infer knowledge.\\
   		 (That's really it.  No prior-knowledge is required:\\
   		 you DERIVE knowledge)
   		 \vspace{12pt}
   \item Definitions of knowledge were a clue:\\
		 Demster-Shafer (generalisation of Bayes)\\
		 Epistemology (Advaita Vedanta wikipedia page)
  \end{itemize}
}

\frame{\frametitle{Advaita Vedanta, Epistemology section}


 \begin{itemize}
 
	\item Pratyakṣa - perception (includes senses, but also "intuition")
   	\item Anumana - inference (where there's smoke there's fire)
   	\item Upamana - comparison, analogy (A is to B as C is to D;\\
   	   		 also included here is the "difference" between two things)
   	\item Arthapatti - postulation, derivation from      
   	      circumstances\\
   	      (Joe is gaining weight; we do not see Joe eat during the day.
   	       Therefore Joe is eating at night)
   	\item Anupalabdi - non-perception, negative/cognitive proof\\
   	      ("there is no jug in this room")
   	\item Sabda - relying on word, testimony of past/present experts
  \end{itemize}
  \bf{ Reverse-Engineers develop these knowledge-derivation skills
  without knowing that they have them! It's incredible and valuable!}

}

\frame{\frametitle{Why do products need reverse-engineering?}
\vspace{9pt}
 \begin{itemize}
   \item The profit-maximising Corporation can't be bothered to provide
   		 documentation or source code
   		\vspace{4pt}
   \item The profit-maximising Corporation is based in China and is
   		 happy to blatantly disregard Copyright law.
   		\vspace{4pt}
   \item The profit-maximising Corporation could be bothered but has
   	 	 realised that they make more money through entrapment of end-users
   		\vspace{4pt}
   \item Bottom line: helping such Corporations helps keep their products
   		 in circulation.
         \vspace{4pt}
  \end{itemize}
  \bf{ Reverse-Engineers by applying their amazing skills actively support
       unethical and pathological Corporations to do harm to end-users
       and to the environment}

}



\frame{\frametitle{How about an alternative?}


 \begin{itemize}
   \item You have unbelievably empowering and powerful skills, far
         beyond those of an average programmer!
   \item Instead of supporting unethical Corporations, why not support
         yourselves?
   \item Transition from full-time to part-time (or work evenings)
   \item How about creating your own products?  (You're a Reverse-Engineer:
         you know how products work, and what you don't know, you already
         know you can to find out!)
   \item How about designing a product and put it on Crowdsupply?
   \item The internet exists: you can find others to team up with,
         in an area of technology that interests you.
  \end{itemize}
  
  \bf{Ultimately you could do a huge amount of good.  With your skill
      there is nothing that can stop you except yourself}
      
}


\end{document}

